Holomorphic field theories and higher algebra

### 2024-10-06 draft.tex
more ref stuff

The reader may (and should) be concerned with the apparent `non-local' form of the action functional in \eqref{eqn:scalar1}.
This can be made rigorous by cohomologically resolving the condition that $\alpha$ be $\del$-closed, rather than strictly imposing the constraint \cite{CLbcov1}.

@@

Any theory of gravity involves, in part, metric structure on a given manifold,
so that solutions to the equations of motion produce metrics satisfying a PDE. 
For a perturbative gravity theory,
one studies deformations of some metric structure on a manifold.
There is a holomorphic avatar of this situation: deformations of complex structure. 
The first appearance of such a theory was motivated by topological string theory and involves complex structures on Calabi--Yau threefolds~\cite{BCOV}.
It is known \textit{Kodaira--Spencer gravity} or the Kodaira-Spencer field theory,
due to its relationship with deformation of complex structures.\footnote{Kodaira and Spencer's work was an important initial clue in the relationship between deformation theory and dg Lie algebras.}
It has been pursued mathematically by Costello and Li \cite{CLbcov1,CLbcov2,LiThesis} as an approach to the quantum higher genus $B$-model.
In these works, Costello and Li also provide a generalization of the theory that makes sense in any complex dimension; 
here we focus just on the most familiar three-dimensional version.
%The fields consist of pairs $(\mu, \eta)$ where
%\begin{align*}
%\mu & \in \Omega^{0,1}(X, \T_X) \\
%\nu & \in \Omega^{0,n-2}(X, K_X \otimes \T^*_X) .
%\end{align*}
%The action functional is
%\beqn
%\int_X \left(\eta \dbar \mu + \frac12 \eta [\mu,\mu]\right)
%\eeqn
%where $[\mu, \mu] \in \Omega^{0,2} (X, \T_X)$ stands for the Lie bracket of holomorphic vector fields extended to the Dolbeault valued sections of $\T_X$ in the natural way.

@@

By a {\em supercharge}, one means an odd supertranslation $\cQ \in \ft$, and
a {\em twist} is a (nonzero) supercharge $\cQ$ such that $[\cQ,\cQ] = \cQ^2 = 0$. 
The classification of twisting supercharges is a completely algebraic question:
see~\autocite{ESsusy} for a complete classification in dimensions from $1$ to~$10$.

@@

If a classical field theory $\cT$ has $\ft$ as a symmetry, 
then a choice of twist $\cQ$ determines a deformation of the theory that we call a {\em twisted supersymmetric theory} $\cT^\cQ$.
If $S$ denotes the action functional of $\cT$, then the action functional $S^\cQ$ of $\cT^\cQ$ has the form
\[
S^\cQ(\varphi) = S(\varphi) + \int_{\RR^d} \varphi  \left(\cQ \cdot \varphi \right) + \cdots 
\]
where the deformation arises from how $\cQ$ acts on the theory 
(the $\cdots$ leaves room for terms non-linear in $\cQ$).
For a systematic discussion of how the twisted theory is defined, 
we refer to~\autocite{CostelloHolomorphic, ESW}.

@@

This theorem applies to many well-known examples from physics:
\begin{itemize}
\item The topological $B$-model of maps from an oriented surface into a Calabi-Yau manifold $X$ (constructed by Li and Li \cite{LiLi} in a setting where \cite{EllSaf} applies) yields an $E_2$ algebra. It is quasi-isomorphic to the well-known Gerstenhaber algebra of polyvector fields on~$X$.
\item The Chern-Simons gauge theory of flat $G$-bundles on an oriented 3-manifold (constructed, essentially, by Axelrod-Singer \cite{AxeSin} and Kontsevich \cite{KonECM} but articulated in this setting by Costello \cite{CosBook}) yields an $E_3$ algebra, which encodes the quantum group $U_\hbar \fg$ under a version of Koszul duality.
\item The Rozansky-Witten theory of maps from an oriented 3-manifold into a complex manifold \cite{ChanLeungLi} yields an $E_3$ algebra \owen{on what?}
\item The Kapustin-Witten gauge theory for oriented 4-manifolds \cite{EGW} yields an $E_4$ algebra, related to deforming functions on the coadjoint quotient stack~$\fg^*/G$.
\end{itemize}
It applies to other examples of topological AKSZ theories, such as BF theories and the Poisson $\sigma$-model.
In all these examples, the underlying elliptic complex appearing in the field theory is a version of the de Rham complex,
so that the chain homotopy $\eta$ is easily constructed by contracting a vector field with differential forms.

@@

For $M = \RR$, we have already discussed how locally constant factorization algebras relate to associative algebras, starting with Example~\ref{ex:associativealgebra}.
Lurie has shown the following vast extension of this example. (See section 5.4.5 of \cite{LurieHA}, particularly Theorem 5.4.5.9.)

@@

This theorem provides an elegant interpretation of BV quantization as a kind of deformation quantization. 
In the setting of mechanics (or one-dimensional field theory, since a particle has a worldline), deformation quantization explains the transition from classical to quantum as deforming the observables from a Poisson algebra to an associative algebra. 
This theorem allows one to interpret quantization of field theories---on manifolds of arbitrary dimension---as a deformation quantization from a 1-shifted Poisson factorization algebra to a plain factorization algebra. 
In fact, in \cite{GLL,Si Kai and Zhengping} it is shown how BV quantization for one-dimensional quantum field theories connects with the ordinary theory of deformation quantization.

@@

Another special case is the map $f: X \to \star$.
Then $f_* \cF = \int_X \cF$ is the global sections of the factorization algebra $\cF$ along $X$.
In the case that $\cF$ is a locally constant factorization algebra, 
this computation agrees with the factorization homology of $\cF$ along $X$, as defined by Ayala-Francis and Lurie \cite{}.
In the case that $\cF$ is a holomorphic factorization algebra on a Riemann surface $\Sigma$, 
the global sections $\int_\Sigma \cF$ is closely related to the chiral homology (or conformal blocks) of the chiral (or vertex) algebra associated to $\cF$ along $\Sigma$.
\owen{I want to assert "(It is the linear dual.)"}

@@

Costello, inspired by the methods of Axelrod-Singer and Kontsevich,
found a framework for perturbative renormalization that works with a large class of elliptic complexes.
When applied to Dolbeault complexes arising in Kodaira-Spencer (or BCOV) theory, 
he and Li found some remarkable simplifications and good behavior \cite{CLbcov}.
In \cite{LiVertex} Li showed, by clever analysis, that holomorphic field theories in complex dimension one (i.e., chiral conformal field theories) are finite to all orders in perturbation theory; 
this result was later extended to arbitrary Riemann surfaces in \cite{LiZhou}.
Following up (cf. his thesis \cite{BWthesis} and its offshoot \cite{Wrenorm}), Williams laid a systematic mathematical foundations:
\begin{itemize}
\item he established key analytic results about holomorphic renormalization, 
showing that it is highly manageable, with no counterterms to 1-loop; and
\item characterized the one-loop anomalies  (i.e., obstructions to BV quantization).
\end{itemize}
These apply to all holomorphic theories on~$\CC^n$,
but the techniques typically extend to more general complex manifolds.

@@

For holomorphic field theories (and generally topological-holomorphic theories, see \cite{WWthf}) however, something remarkable happens.

@@

In other words, divergences do not appear when constructing the perturbative quantization of {\em any} holomorphic field theory on~$\CC^d$.

@@

Since renormalization is so well-behaved, however, such anomalies are amenable to explicit characterization.
For example, at one-loop in perturbation theory, an explicit formula for such anomalies for holomorphic theories on $\C^d$ is given in \cite{Whol}.
It involves a sum over Feynman weights $W_\Gamma$ where $\Gamma$ is a polygon with $(d+1)$~vertices.

>>>>

The reader may (and should) be concerned with the apparent ``non-local'' form of the action functional in \eqref{eqn:scalar1}.
This can be made rigorous by cohomologically resolving the condition that $\alpha$ be $\del$-closed, rather than strictly imposing the constraint~\cite{CLbcov}.

@@

Any theory of gravity involves, in part, metric structure on a given manifold,
so that solutions to the equations of motion produce metrics satisfying a PDE. 
For a perturbative gravity theory,
one studies deformations of some metric structure on a manifold.
There is a holomorphic avatar of this situation: deformations of complex structure. 
The first appearance of such a theory was motivated by topological string theory and involves complex structures on Calabi--Yau threefolds~\cite{BCOV}.
It is known \textit{Kodaira--Spencer gravity} or the Kodaira-Spencer field theory,
due to its relationship with deformation of complex structures.\footnote{Kodaira and Spencer's work was an important initial clue in the relationship between deformation theory and dg Lie algebras.}
It has been pursued mathematically by Costello and Li as an approach to the quantum higher genus $B$-model, starting with \cite{CLbcov} and continuing in a series of papers.
In these works, Costello and Li also provide a generalization of the theory that makes sense in any complex dimension; 
here we focus just on the most familiar three-dimensional version.
%The fields consist of pairs $(\mu, \eta)$ where
%\begin{align*}
%\mu & \in \Omega^{0,1}(X, \T_X) \\
%\nu & \in \Omega^{0,n-2}(X, K_X \otimes \T^*_X) .
%\end{align*}
%The action functional is
%\beqn
%\int_X \left(\eta \dbar \mu + \frac12 \eta [\mu,\mu]\right)
%\eeqn
%where $[\mu, \mu] \in \Omega^{0,2} (X, \T_X)$ stands for the Lie bracket of holomorphic vector fields extended to the Dolbeault valued sections of $\T_X$ in the natural way.

@@

By a {\em supercharge}, one means an odd supertranslation $\cQ \in \ft$, and
a {\em twist} is a (nonzero) supercharge $\cQ$ such that $[\cQ,\cQ] = \cQ^2 = 0$. 
The classification of twisting supercharges is a completely algebraic question:
see~\autocite{EllSaf} for a complete classification in dimensions from $1$ to~$10$.

@@

If a classical field theory $\cT$ has $\ft$ as a symmetry, 
then a choice of twist $\cQ$ determines a deformation of the theory that we call a {\em twisted supersymmetric theory} $\cT^\cQ$.
If $S$ denotes the action functional of $\cT$, then the action functional $S^\cQ$ of $\cT^\cQ$ has the form
\[
S^\cQ(\varphi) = S(\varphi) + \int_{\RR^d} \varphi  \left(\cQ \cdot \varphi \right) + \cdots 
\]
where the deformation arises from how $\cQ$ acts on the theory 
(the $\cdots$ leaves room for terms non-linear in $\cQ$).
For a systematic discussion of how the twisted theory is defined, 
we refer to~\autocite{CosSUSY, ESW}.

@@

This theorem applies to many well-known examples from physics:
\begin{itemize}
\item The topological $B$-model of maps from an oriented surface into a Calabi-Yau manifold $X$ yields an $E_2$ algebra. It is quasi-isomorphic to the well-known Gerstenhaber algebra of polyvector fields on~$X$. (The quantization was constructed by Li and Li \cite{LiLi} in a setting where \cite{EllSaf} applies.)
\item The Chern-Simons gauge theory of flat $G$-bundles on an oriented 3-manifold yields an $E_3$ algebra, which encodes the quantum group $U_\hbar \fg$ under a version of Koszul duality. 
(The quantization was constructed, essentially, by Axelrod-Singer \cite{AxeSing1,AxeSing2} and Kontsevich \cite{KonECM} but articulated in this setting by Costello~\cite{CosBook}.) 
\item The Rozansky-Witten theory of maps from an oriented 3-manifold into a complex manifold yields an $E_3$ algebra. (The quantization was constructed by Chan, Leung, and Li~\cite{ChanLeungLi}.)
\item The Kapustin-Witten gauge theory for oriented 4-manifolds yields an $E_4$ algebra, related to deforming functions on the coadjoint quotient stack~$\fg^*/G$, (The quantization was constructed in~\cite{EGW}.)
\end{itemize}
It applies to other examples of topological AKSZ theories, such as BF theories and the Poisson $\sigma$-model.
In all these examples, the underlying elliptic complex appearing in the field theory is a version of the de Rham complex,
so that the chain homotopy $\eta$ is easily constructed by contracting a vector field with differential forms.

@@

For $M = \RR$, we have already discussed how locally constant factorization algebras relate to associative algebras, starting with Example~\ref{ex:associativealgebra}.
Lurie has shown the following vast extension of this example. (See section 5.4.5 of \cite{LurieHA}, particularly Theorem~5.4.5.9.)

@@

This theorem provides an elegant interpretation of BV quantization as a kind of deformation quantization. 
In the setting of mechanics (or one-dimensional field theory, since a particle has a worldline), deformation quantization explains the transition from classical to quantum as deforming the observables from a Poisson algebra to an associative algebra. 
This theorem allows one to interpret quantization of field theories---on manifolds of arbitrary dimension---as a deformation quantization from a 1-shifted Poisson factorization algebra to a plain factorization algebra. 
In fact, in \cite{GLL,GLX} it is shown how BV quantization for one-dimensional quantum field theories connects with the ordinary theory of deformation quantization.

@@

Another special case is the map $f: X \to \star$.
Then $f_* \cF = \int_X \cF$ is the global sections of the factorization algebra $\cF$ along $X$.
In the case that $\cF$ is a locally constant factorization algebra, 
this computation agrees with the factorization homology of $\cF$ along $X$, as defined by Ayala-Francis and Lurie.
In the case that $\cF$ is a holomorphic factorization algebra on a Riemann surface $\Sigma$, 
the global sections $\int_\Sigma \cF$ is closely related to the chiral homology (or conformal blocks) of the chiral (or vertex) algebra associated to $\cF$ along $\Sigma$.
\owen{I want to assert "(It is the linear dual.)"}

@@

Costello, inspired by the methods of Axelrod-Singer and Kontsevich,
found a framework for perturbative renormalization that works with a large class of elliptic complexes.
When applied to Dolbeault complexes arising in Kodaira-Spencer (or BCOV) theory, 
he and Li found some remarkable simplifications and good behavior \cite{CLbcov}.
In \cite{LiVertex} Li showed, by clever analysis, that holomorphic field theories in complex dimension one (i.e., chiral conformal field theories) are finite to all orders in perturbation theory; 
this result was later extended to arbitrary Riemann surfaces in \cite{LiZhou}.
Following up (cf. his thesis \cite{Wthesis} and its offshoot \cite{Wrenorm}), Williams laid a systematic mathematical foundations:
\begin{itemize}
\item he established key analytic results about holomorphic renormalization, 
showing that it is highly manageable, with no counterterms to 1-loop; and
\item characterized the one-loop anomalies  (i.e., obstructions to BV quantization).
\end{itemize}
These apply to all holomorphic theories on~$\CC^n$,
but the techniques typically extend to more general complex manifolds.

@@

For holomorphic field theories, however, something remarkable happens.

@@

In other words, divergences do not appear when constructing the perturbative quantization of {\em any} holomorphic field theory on~$\CC^d$.
(Similar results hold for topological-holomorphic theories~\cite{WWthf}.)

@@

Since renormalization is so well-behaved, however, such anomalies are amenable to explicit characterization.
For example, at one-loop in perturbation theory, an explicit formula for such anomalies for holomorphic theories on $\C^d$ is given in \cite{Wrenorm}.
It involves a sum over Feynman weights $W_\Gamma$ where $\Gamma$ is a polygon with $(d+1)$~vertices.
